---
title: "SP21_boxplot"
author: "Dr. Joy Buongiorno"
output:
  pdf_document: default
  html_document: default
goal: Vizualize transcription trends of different copies of the SP21 gene with depth
  across 4 sites and place statistics on depth trends.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load libraries}
library(ggplot2)
library(dplyr)
library(agricolae)
library(tidyverse)
library(ggpubr)
```

```{r Upload data}
sp21<-read.csv("SP21_coverage_forR.csv")
```

```{r initial boxplot that shows trends}
sp21box<-ggboxplot(sp21, x="Depth", y = "Normalized_coverage", 
                color="Site", palette=c("Dark2"),
                add="jitter", 
                ylab="Normalized coverage",
               yscale="log10")
facet(sp21box, facet.by= "Site")
```

```{Prepare data for library-specific testing of normality}
Sp21_F01<- filter(sp21, Site == "F", Depth =="0.5")
Sp21_F12<- filter(sp21, Site =="F", Depth =="1.5")
Sp21_F23<- filter(sp21, Site =="F", Depth =="2.5")
Sp21_F34<- filter(sp21, Site == "F", Depth =="3.5")
Sp21_F45<- filter(sp21, Site =="F", Depth =="4.5")
Sp21_F56<- filter(sp21, Site =="F", Depth =="5.5")

Sp21_P01<- filter(sp21, Site =="P", Depth =="0.5")
Sp21_P01<- Sp21_P01[-c(5), ] #Zero coverage
Sp21_P12<- filter(sp21, Site =="P", Depth =="1.5")
Sp21_P23<- filter(sp21, Site =="P", Depth == "2.5")
Sp21_P34<- filter(sp21, Site =="P", Depth =="3.5")
Sp21_P56<- filter(sp21, Site =="P", Depth =="5.5")

Sp21_AB05<- filter(sp21, Site =="AB", Depth =="0.5")
Sp21_AB05<- Sp21_AB05[-c(5), ] #Zero coverage
Sp21_AB12<- filter(sp21, Site =="AB", Depth == "1.5")
Sp21_AB23<- filter(sp21, Site =="AB", Depth =="2.5")
Sp21_AB34<- filter(sp21, Site =="AB", Depth =="3.5")

Sp21_AC12<- filter(sp21, Site =="AC", Depth =="1.5")
Sp21_AC12<- Sp21_AC12[-c(5), ] #Zero coverage
Sp21_AC23<- filter(sp21, Site =="AC", Depth =="2.5")
Sp21_AC34<- filter(sp21, Site =="AC", Depth =="3.5")
Sp21_AC45<- filter(sp21, Site =="AC", Depth == "4.5")
```

Normality testing. At alpha = 0.05 in a Shapiro-Wilks test, we reject the null hypothesis that our distribution is normal. Therefore, when p > 0.05 in the shapiro test, the data is normally-distributed. 

```{Shapiro test for normality}
shapiro.test(log10(Sp21_F01$Normalized_coverage)) #log-transformed is normal. p-value = 0.3228
qqnorm(log10(Sp21_F01$Normalized_coverage))

shapiro.test(log10(Sp21_F12$Normalized_coverage)) #log-transformed is normal. p-value = 0.7928
qqnorm(log10(Sp21_F12$Normalized_coverage))

#this library may need to be removed from this analysis -- small
shapiro.test(log10(Sp21_F23$Normalized_coverage)) #log-transformed is normal. p-value = 0.1065
qqnorm(log10(Sp21_F23$Normalized_coverage))

shapiro.test(log10(Sp21_F34$Normalized_coverage))  #log-transformed is normal. p-value =  0.1215
qqnorm(log10(Sp21_F34$Normalized_coverage))

#this library may need to be removed from this analysis -- small
shapiro.test(log10(Sp21_F45$Normalized_coverage)) #log-transformed is normal. p-value = 0.1278
qqnorm(log10(Sp21_F45$Normalized_coverage))

shapiro.test(log10(Sp21_F56$Normalized_coverage)) #log-transformed is normal. p-value = 0.2168
qqnorm(log10(Sp21_F56$Normalized_coverage))

shapiro.test(log10(Sp21_P01$Normalized_coverage)) #log-transformed is normal. p-value = 0.0544
qqnorm(Sp21_P01$Normalized_coverage)

shapiro.test((log10(Sp21_P12$Normalized_coverage))) #log-transformed is normal. p-value = 0.659
qqnorm(Sp21_P12$Normalized_coverage)

shapiro.test(log10(Sp21_P23$Normalized_coverage)) #log-transformed is normal. p-value = 0.5738
qqnorm(log10(Sp21_P23$Normalized_coverage))

shapiro.test(log10(Sp21_P34$Normalized_coverage))  #log-transformed is normal. p-value = 0.4896
qqnorm(log10(Sp21_P34$Normalized_coverage))

shapiro.test(log10(Sp21_P56$Normalized_coverage)) #log-transformed is normal. p-value = 0.06494
qqnorm(log10(Sp21_P56$Normalized_coverage))

shapiro.test(log10(Sp21_AB05$Normalized_coverage)) #log-transformed is normal.  p-value = 0.2125
qqnorm(log10(Sp21_AB05$Normalized_coverage))

shapiro.test(log10(Sp21_AB12$Normalized_coverage)) #log-transformed is normal. p-value = 0.9919
qqnorm(log10(Sp21_AB12$Normalized_coverage))

shapiro.test(log10(Sp21_AB23$Normalized_coverage))  #log-transformed is normal. p-value = 0.9231
qqnorm(log10(Sp21_AB23$Normalized_coverage))

shapiro.test(log10(Sp21_AB34$Normalized_coverage)) #log-transformed is normal. p-value = 0.8091
qqnorm(log10(Sp21_AB34$Normalized_coverage))

shapiro.test(log10(Sp21_AC12$Normalized_coverage)) #log-transformed is normal. p-value = 0.4367
qqnorm(log10(Sp21_AC12$Normalized_coverage))

shapiro.test(log10(Sp21_AC23$Normalized_coverage)) #log-transformed is normal. p-value = 0.2606
qqnorm(log10(Sp21_AC23$Normalized_coverage))

shapiro.test(log10(Sp21_AC34$Normalized_coverage)) #log-transformed is normal. p-value = 0.788
qqnorm(log10(Sp21_AC34$Normalized_coverage))

shapiro.test(log10(Sp21_AC45$Normalized_coverage)) #log-transformed is normal. p-value = 0.3404
qqnorm(log10(Sp21_AC45$Normalized_coverage))
```

Now that we know all of our values are normally-distributed when log-transformed (which is how we present our data), we can move forwrad with Tukey means testing.

```{r Site F}
sp21_F<- filter(sp21, Site == "F")
sp21_F$Depth<-as.factor(sp21_F$Depth)
sp21_F$log<-log10(sp21_F$Normalized_coverage)
sp21_F_aov<-aov(log ~ Depth, data = sp21_F)
summary(sp21_F_aov)

#Get letter designations
library(agricolae)
sp21_F_tukey_groups<-HSD.test(sp21_F_aov, "Depth", group = TRUE) #all depths are "a"; no difference between means

#Get stats with p-values
sp21_F_tukey_stats<-TukeyHSD(sp21_F_aov, which ="Depth")
plot(sp21_F_tukey_stats)

#Extract the letters
res<-sp21_F_tukey_groups$groups
res$Depth<-rownames(res)

#Place letters on dataframe
sp21_F<-merge(x=sp21_F,y=res[ , c("groups", "Depth")], by="Depth")
```

```{r Site P}
sp21_P<- filter(sp21, Site == "P")
sp21_P$Depth<-as.factor(sp21_P$Depth)
sp21_P$log<-log10(sp21_P$Normalized_coverage)
sp21_P<-sp21_P[-c(5), ] #remove infinity value
sp21_P_aov<-aov(log ~ Depth, data = sp21_P)
summary(sp21_P_aov)

#Get letter designations
library(agricolae)
sp21_P_tukey_groups<-HSD.test(sp21_P_aov, "Depth", group = TRUE) #levels "a", "ab" and "b"; no difference between means

#Get stats with p-values
sp21_P_tukey_stats<-TukeyHSD(sp21_P_aov, which ="Depth")
plot(sp21_P_tukey_stats)

#Extract the letters
res2<-sp21_P_tukey_groups$groups
res2$Depth<-rownames(res2)

#Place letters on dataframe
sp21_P<-merge(x=sp21_P,y=res2[ , c("groups", "Depth")], by="Depth")
```

```{r site AB}
sp21_AB<- filter(sp21, Site == "AB")
sp21_AB$Depth<-as.factor(sp21_AB$Depth)
sp21_AB$log<-log10(sp21_AB$Normalized_coverage)
sp21_AB<-sp21_AB[-c(5), ] #remove infinity value
sp21_AB_aov<-aov(log ~ Depth, data = sp21_AB)
summary(sp21_AB_aov)

#Get letter designations
library(agricolae)
sp21_AB_tukey_groups<-HSD.test(sp21_AB_aov, "Depth", group = TRUE) #levels are "a", "ab", and "b" ; no difference between means

#Get stats with p-values
sp21_AB_tukey_stats<-TukeyHSD(sp21_AB_aov, which ="Depth")
plot(sp21_AB_tukey_stats)

#Extract the letters
res3<-sp21_AB_tukey_groups$groups
res3$Depth<-rownames(res3)

#Place letters on dataframe
sp21_AB<-merge(x=sp21_AB,y=res3[ , c("groups", "Depth")], by="Depth")
```

```{r site AC}
sp21_AC<- filter(sp21, Site == "AC")
sp21_AC$Depth<-as.factor(sp21_AC$Depth)
sp21_AC$log<-log10(sp21_AC$Normalized_coverage)
sp21_AC<-sp21_AC[-c(5), ] #remove infinity value
sp21_AC_aov<-aov(log ~ Depth, data = sp21_AC)
summary(sp21_AC_aov)

#Get letter designations
library(agricolae)
sp21_AC_tukey_groups<-HSD.test(sp21_AC_aov, "Depth", group = TRUE) #all depths are "a" or "ab"; no difference between means

#Get stats with p-values
sp21_AC_tukey_stats<-TukeyHSD(sp21_AC_aov, which ="Depth")
plot(sp21_AC_tukey_stats)

#Extract the letters
res4<-sp21_AC_tukey_groups$groups
res4$Depth<-rownames(res4)

#Place letters on dataframe
sp21_AC<-merge(x=sp21_AC,y=res4[ , c("groups", "Depth")], by="Depth")
```

Change the letters in each data frame so that each site has unique identifier for plotting different colors in each panel.

```{r new unique letters}
sp21_F$groups<-gsub('a', 'a (site F)', sp21_F$groups)

sp21_P$groups<-gsub("\\<a\\>", 'a (site P)', sp21_P$groups)
sp21_P$groups<-gsub("\\<ab\\>", 'ab (site P)', sp21_P$groups)
sp21_P$groups<-gsub("\\<b\\>", 'b (site P)', sp21_P$groups)

sp21_AB$groups<-gsub("\\<a\\>", 'a (site AB)', sp21_AB$groups)
sp21_AB$groups<-gsub("\\<ab\\>", 'ab (site AB)', sp21_AB$groups)
sp21_AB$groups<-gsub("\\<b\\>", 'b (site AB)', sp21_AB$groups)

sp21_AC$groups<-gsub('a', 'a (site AC)', sp21_AC$groups)
```

```{r paste together, make palette, and plot}
sp21_all<-rbind(sp21_F, sp21_P, sp21_AB, sp21_AC)
names(sp21_all)[names(sp21_all) == "groups"] <- "intrasite_groups"

#Design the palette with strategically-selected colors that are color-blind friendly.
tukey_pal<-c("a (site F)" = "#84158e", "a (site P)"  = "#03396c", "ab (site P)"="#6497b1", "b (site P)"= "#b3cde0", "a (site AB)" ="#4B3E40", "ab (site AB)" = "#E4D7D9", "b (site AB)" = "#FDF0F2", "a (site AC)" = "#bdd6bf" )

sp21_all$TPM<-sp21_all$Normalized_coverage*1000000
sp21box_tukey<-ggboxplot(sp21_all, x="Depth", y = "TPM", 
                   fill="intrasite_groups", palette=tukey_pal,
                   add="jitter", 
                   ylab="log(TPM)",
                   xlab="Depth (cmbsf)",
                   yscale="log10")
facet(sp21box_tukey, facet.by= "Site") 
```
